---
title: "Pairwise and Marginal distribution"
author: "Shihui Zhu"
date: 2021-11-28
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(patchwork)
```

```{r in, message=FALSE, warning=FALSE}
cdi = read_csv("./cdi.csv") %>% 
  janitor::clean_names() %>%
  mutate(
    crime_rate = crimes/pop * 1000
  )
```

## Normal?

```{r, echo=FALSE, message=FALSE}
p1 <- cdi %>%
  ggplot(aes(x = totalinc)) +
  geom_histogram()
p2 <- cdi %>%
  ggplot(aes(x = pcincome)) +
  geom_histogram()
p3 <- cdi %>%
  ggplot(aes(x = unemp)) +
  geom_histogram()
p4 <- cdi %>%
  ggplot(aes(x = poverty)) +
  geom_histogram()
p5 <- cdi %>%
  ggplot(aes(x = bagrad)) +
  geom_histogram()
p6 <- cdi %>%
  ggplot(aes(x = hsgrad)) +
  geom_histogram()

(p1+p2+p3) / (p4+p5+p6)
```

Note that the plots are not normal. They are all skewed. 

## Marginal Distribution

### Total personal income
```{r}
#shapiro.test(cdi$totalinc)

# Scatter plot with regression line overlaid and 95% confidence bands
cdi %>% 
  ggplot(aes(totalinc, crime_rate)) + geom_point(color='blue') + theme_bw(base_size=20) +
  geom_smooth(method='lm', se=TRUE, color='red') +
  labs(x="Total personal income", y="Crimes rate")

# Simple linear regression 
reg_totalinc = lm(crime_rate ~ totalinc, data = cdi)
summary(reg_totalinc)
```
There is a significant linear increasing between crime rate and total personal income


### Per capita income

```{r}
# Scatter plot with regression line overlaid and 95% confidence bands
cdi %>% 
  ggplot(aes(pcincome, crime_rate)) + geom_point(color='blue') + theme_bw(base_size=20) +
  geom_smooth(method='lm', se=TRUE, color='red') +
  labs(x="Per capita income", y="Crimes rate")

# Simple linear regression 
reg_pcincome = lm(crime_rate ~ pcincome, data = cdi)
summary(reg_pcincome)
```

There is NO significant linear relationship between crime rate and per capita income.

### Percent unemployment

```{r}
# Scatter plot with regression line overlaid and 95% confidence bands
cdi %>% 
  ggplot(aes(unemp, crime_rate)) + geom_point(color='blue') + theme_bw(base_size=20) +
  geom_smooth(method='lm', se=TRUE, color='red') +
  labs(x="Percent unemployment", y="Crimes rate")

# Simple linear regression 
reg_unemp = lm(crime_rate ~ unemp, data = cdi)
summary(reg_unemp)
```
There is NO significant linear relationship between crime rate and percent unemployment.

### Percent below poverty level

```{r}
# Scatter plot with regression line overlaid and 95% confidence bands
cdi %>% 
  ggplot(aes(poverty, crime_rate)) + geom_point(color='blue') + theme_bw(base_size=20) +
  geom_smooth(method='lm', se=TRUE, color='red') +
  labs(x="Percent below poverty level", y="Crimes rate")

# Simple linear regression 
reg_poverty = lm(crime_rate ~ poverty, data = cdi)
summary(reg_poverty)
```
There is significant linear increasing between crime rate and percent unemployment.

### Percent bachelor’s degrees

```{r}
# Scatter plot with regression line overlaid and 95% confidence bands
cdi %>% 
  ggplot(aes(bagrad, crime_rate)) + geom_point(color='blue') + theme_bw(base_size=20) +
  geom_smooth(method='lm', se=TRUE, color='red') +
  labs(x="Percent bachelor’s degrees", y="Crimes rate")

# Simple linear regression 
reg_bagrad = lm(crime_rate ~ bagrad, data = cdi)
summary(reg_bagrad)
```
There is NO significant linear relationship between crime rate and percent bachelor’s degrees.

### Percent high school graduates

```{r}
# Scatter plot with regression line overlaid and 95% confidence bands
cdi %>% 
  ggplot(aes(hsgrad, crime_rate)) + geom_point(color='blue') + theme_bw(base_size=20) +
  geom_smooth(method='lm', se=TRUE, color='red') +
  labs(x="Percent high school graduates", y="Crimes rate")

# Simple linear regression 
reg_hsgrad = lm(crime_rate ~ hsgrad, data = cdi)
summary(reg_hsgrad)
```
There is significant linear decreasing between crime rate and percent unemployment.

### Summary

* There is no linear relationship between per capita income, percent unemployment, and percent bachelor's degrees with crimes rate. 

We then want to examine whether there is linear relationship between each variables. 

## Pairwise comparsion

### Percent high school graduates v.s. Total personal income

```{r}
# Scatter plot with regression line overlaid and 95% confidence bands
cdi %>% 
  ggplot(aes(hsgrad, totalinc)) + geom_point(color='blue') + theme_bw(base_size=20) +
  geom_smooth(method='lm', se=TRUE, color='red') +
  labs(x="Percent high school graduates", y="Total personal income")

# Simple linear regression 
hs_to = lm(totalinc ~ hsgrad, data = cdi)
summary(hs_to)
```
There is NO linear relationship between total personal income and percent high school graduates.

### Poverty v.s. Percent high school graduates

```{r}
# Scatter plot with regression line overlaid and 95% confidence bands
cdi %>% 
  ggplot(aes(poverty, hsgrad)) + geom_point(color='blue') + theme_bw(base_size=20) +
  geom_smooth(method='lm', se=TRUE, color='red') +
  labs(x="Percent below poverty level", y="Percent high school graduates")

# Simple linear regression 
po_hs = lm(hsgrad ~ poverty, data = cdi)
summary(po_hs)
```
There is a significant linear decreasing between percent below poverty level and percent high school graduates.

### Poverty v.s. Total personal income

```{r}
# Scatter plot with regression line overlaid and 95% confidence bands
cdi %>% 
  ggplot(aes(poverty, totalinc)) + geom_point(color='blue') + theme_bw(base_size=20) +
  geom_smooth(method='lm', se=TRUE, color='red') +
  labs(x="Percent below poverty level", y="Total personal income")

# Simple linear regression 
po_to = lm(totalinc ~ poverty, data = cdi)
summary(po_to)
```
There is NO linear relationship between total personal income and percent below poverty level.

**With other variables**

### Percent unemployment v.s. Percent high school graduates

```{r}
# Scatter plot with regression line overlaid and 95% confidence bands
cdi %>% 
  ggplot(aes(unemp, hsgrad)) + geom_point(color='blue') + theme_bw(base_size=20) +
  geom_smooth(method='lm', se=TRUE, color='red') +
  labs(x="Percent unemployment", y="Percent high school graduates")

# Simple linear regression 
hs_un = lm(hsgrad ~ unemp, data = cdi)
summary(hs_un)
```
There is a significant linear decreasing between percent unemployment and percent high school graduates.

### Per capita income v.s. Percent high school graduates

```{r}
# Scatter plot with regression line overlaid and 95% confidence bands
cdi %>% 
  ggplot(aes(pcincome, hsgrad)) + geom_point(color='blue') + theme_bw(base_size=20) +
  geom_smooth(method='lm', se=TRUE, color='red') +
  labs(x="Per capita income", y="Percent high school graduates")

# Simple linear regression 
hs_pc = lm(hsgrad ~ pcincome, data = cdi)
summary(hs_pc)
```
There is a significant linear increasing between per capita income and percent high school graduates.

### Percent bachelor’s degrees v.s. Percent high school graduates

```{r}
# Scatter plot with regression line overlaid and 95% confidence bands
cdi %>% 
  ggplot(aes(hsgrad, bagrad)) + geom_point(color='blue') + theme_bw(base_size=20) +
  geom_smooth(method='lm', se=TRUE, color='red') +
  labs(x="Percent high school graduates", y="Percent bachelor’s degrees")

# Simple linear regression 
ba_hs = lm(bagrad ~ hsgrad, data = cdi)
summary(ba_hs)
```
There is a significant linear increasing between percent bachelor’s degrees and percent high school graduates

### Total personal income v.s. Percent bachelor’s degrees

```{r}
# Scatter plot with regression line overlaid and 95% confidence bands
cdi %>% 
  ggplot(aes(totalinc, bagrad)) + geom_point(color='blue') + theme_bw(base_size=20) +
  geom_smooth(method='lm', se=TRUE, color='red') +
  labs(x="Total personal income", y="Percent bachelor’s degrees")

# Simple linear regression 
ba_to = lm(bagrad ~ totalinc, data = cdi)
summary(ba_to)
```

There is a significant linear increasing between percent bachelor’s degrees and total personal income.

### Total personal income v.s. Per capita income

```{r}
# Scatter plot with regression line overlaid and 95% confidence bands
cdi %>% 
  ggplot(aes(totalinc, pcincome)) + geom_point(color='blue') + theme_bw(base_size=20) +
  geom_smooth(method='lm', se=TRUE, color='red') +
  labs(x="Total personal income", y="Per capita income")

# Simple linear regression 
to_pc = lm(pcincome ~ totalinc, data = cdi)
summary(to_pc)
```
There is a significant linear increasing between per capita income and total personal income.

### Total personal income v.s. percent unemployment

```{r}
# Scatter plot with regression line overlaid and 95% confidence bands
cdi %>% 
  ggplot(aes(totalinc, unemp)) + geom_point(color='blue') + theme_bw(base_size=20) +
  geom_smooth(method='lm', se=TRUE, color='red') +
  labs(x="Total personal income", y="Percent unemployment")

# Simple linear regression 
to_un = lm(unemp ~ totalinc, data = cdi)
summary(to_un)
```
There is NO linear relationship between total personal income and percent unemployment.

### Per capita income v.s. percent unemployment

```{r}
# Scatter plot with regression line overlaid and 95% confidence bands
cdi %>% 
  ggplot(aes(pcincome, unemp)) + geom_point(color='blue') + theme_bw(base_size=20) +
  geom_smooth(method='lm', se=TRUE, color='red') +
  labs(x="Per capita income", y="Percent unemployment")

# Simple linear regression 
pc_un = lm(unemp ~ pcincome, data = cdi)
summary(pc_un)
```
There is a significant linear decreasing between per capita income and percent unemployment.

### Per capita income v.s. percent below poverty level

```{r}
# Scatter plot with regression line overlaid and 95% confidence bands
cdi %>% 
  ggplot(aes(pcincome, poverty)) + geom_point(color='blue') + theme_bw(base_size=20) +
  geom_smooth(method='lm', se=TRUE, color='red') +
  labs(x="Per capita income", y="Percent below poverty level")

# Simple linear regression 
pc_po = lm(poverty ~ pcincome, data = cdi)
summary(pc_po)
```
There is a significant linear decreasing between per capita income and percent below poverty level.

### Per capita income v.s. Percent bachelor’s degrees

```{r}
# Scatter plot with regression line overlaid and 95% confidence bands
cdi %>% 
  ggplot(aes(pcincome, bagrad)) + geom_point(color='blue') + theme_bw(base_size=20) +
  geom_smooth(method='lm', se=TRUE, color='red') +
  labs(x="Per capita income", y="Percent bachelor’s degrees")

# Simple linear regression 
ba_pc = lm(bagrad ~ pcincome, data = cdi)
summary(ba_pc)
```

There is a significant linear increasing between per capita income and percent below poverty level.

### Percent unemployment v.s. percent below poverty level

```{r}
# Scatter plot with regression line overlaid and 95% confidence bands
cdi %>% 
  ggplot(aes(unemp, poverty)) + geom_point(color='blue') + theme_bw(base_size=20) +
  geom_smooth(method='lm', se=TRUE, color='red') +
  labs(x="Percent unemployment", y="Percent below poverty level")

# Simple linear regression 
po_un = lm(poverty ~ unemp, data = cdi)
summary(po_un)
```
There is a significant linear increasing between percent unemployment and percent below poverty level.

### Percent unemployment v.s. percent bachelor's degrees

```{r}
# Scatter plot with regression line overlaid and 95% confidence bands
cdi %>% 
  ggplot(aes(unemp, bagrad)) + geom_point(color='blue') + theme_bw(base_size=20) +
  geom_smooth(method='lm', se=TRUE, color='red') +
  labs(x="Percent unemployment", y="Percent bachelor's degrees")

# Simple linear regression 
ba_un = lm(bagrad ~ unemp, data = cdi)
summary(ba_un)
```
There is a significant linear decreasing between per capita income and percent unemployment

### Percent below poverty level v.s. percent bachelor's degrees

```{r}
# Scatter plot with regression line overlaid and 95% confidence bands
cdi %>% 
  ggplot(aes(poverty, bagrad)) + geom_point(color='blue') + theme_bw(base_size=20) +
  geom_smooth(method='lm', se=TRUE, color='red') +
  labs(x="Percent below poverty level", y="Percent bachelor's degrees")

# Simple linear regression 
ba_po = lm(bagrad ~ poverty, data = cdi)
summary(ba_po)
```

There is a significant linear decreasing between percent below poverty level and percent bachelor's degrees

## Summary

* There is linear relationship between total personal income (+), poverty(+), and percent high school graduates(-) with crimes rate.

* There is linear relationship between percent bachelor’s degrees(+), and per capita income(+), with total personal income.

* There is linear relationship between percent high school graduates(-), percent bachelor’s degrees(-), and per capita income(-) with percent below poverty level.

* There is linear relationship between percent below poverty level(-), percent unemployment(-), percent bachelor’s degrees(+), and per capita income(+) with percent high school graduates.
