---
title: "Pairwise"
author: "Anyu Zhu"
date: "11/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(corrplot)
library(GGally)
library(ggplot2)
```

```{r}
cdi <- read_csv("cdi.csv") %>% 
  janitor::clean_names() %>% 
  mutate(state = as.factor(state))

var <- 
  cdi[3:17] %>% 
  select(-crimes)

anyNA(var)
```
No NA in the dataset.

## Pairwise Comparison

```{r}
pairs(var, pch = 20)
ggcorr(data = var, label = TRUE, label_size = 2, hjust = 0.8)
```

From the correlation plot, we can see there are terms with high correlation:
Number of active physicians & Number of hospital beds;
Number of hospital beds & Total personal income;
Number of active physicians & Total personal income;
Total population & Number of active physicians;
Total population & Number of hospital beds;
Total population & Total personal Income

## Interaction Terms

Include interaction terms of highly correlated variables into the dataset.
```{r}
with_inter = var %>% 
  mutate(
    docs_beds = docs*beds,
    beds_totalinc = beds*totalinc,
    docs_totalinc = docs*totalinc,
    pop_docs = pop*docs,
    pop_beds = pop*beds,
    pop_totalinc = pop*totalinc
    )
```

## Variable Selection

