---
title: "Interaction & Diagnostic"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include = FALSE, message = FALSE}
library(tidyverse)
library(dplyr)
library(MASS)
library(ggplot2)
library(GGally)
library(leaps)
library(lattice)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

set.seed(1)
```

## Data upload

```{r, message = FALSE}
crime_df = 
  read_csv("cdi.csv") %>% 
  mutate(
    region = factor(region),
    pop_density = pop / area,
    beds_density = beds / pop,
    docs_density = docs / pop,
    crm_1000 = 1000 * crimes / pop
  ) %>% 
  dplyr::select(-crimes, -id, -area, -pop, -beds, -docs) %>% 
  dplyr::select(crm_1000, everything()) %>%
  drop_na()
```

## EDA
```{r}
crime_state = crime_df %>% 
  group_by(state) %>% 
  summarize(crime_rate = mean(crm_1000))

crime_state[order(-rank(crime_state$crime_rate)),] %>% 
  knitr::kable()
```

## Partial ANOVA model select without interaction

```{r}
#linear regression model with full predictors
fit_no_inte = lm(crm_1000 ~ pop18 +totalinc + poverty + hsgrad + beds_density + docs_density + region + pop_density, data = crime_df)
summary(fit_no_inte)
```

Next using partial ANOVA to test whether large model is superior

### no pop_density
```{r}
fit_1 = lm(crm_1000 ~ pop18 + totalinc + poverty + hsgrad + beds_density + docs_density + region, data = crime_df)
summary(fit_1)
# compare nested (small vs large) models
# Ho: smaller model is defensible
anova(fit_1, fit_no_inte)
```

Reject H0, large model is superior

### no region

```{r}
fit_2 = lm(crm_1000 ~ pop18 + totalinc + poverty + hsgrad + beds_density + docs_density + pop_density, data = crime_df)
summary(fit_2)
# compare nested (small vs large) models
# Ho: smaller model is defensible
anova(fit_2, fit_no_inte)
```

Reject H0, large model is superior


### no docs_density

```{r}
fit_3 = lm(crm_1000 ~  pop18 + totalinc + poverty + hsgrad + beds_density + region + pop_density, data = crime_df)
summary(fit_3)
# compare nested (small vs large) models
# Ho: smaller model is defensible
anova(fit_3, fit_no_inte)
```

Fail to reject H0, large model is not superior, we should keep the smaller one.

### no beds_density

```{r}
fit_4 = lm(crm_1000 ~ pop18 + totalinc + poverty + hsgrad + region + pop_density + docs_density, data = crime_df)
summary(fit_4)
# compare nested (small vs large) models
# Ho: smaller model is defensible
anova(fit_4, fit_no_inte)
```

Reject H0, large model is superior


### no hsgrad
```{r}
fit_5 = lm(crm_1000 ~ pop18 + totalinc + poverty + region + pop_density + docs_density + beds_density, data = crime_df)
summary(fit_5)
# compare nested (small vs large) models
# Ho: smaller model is defensible
anova(fit_5, fit_no_inte)
```

Fail to reject H0, large model is not superior, we should keep the smaller one.


### no poverty
```{r}
fit_6 = lm(crm_1000 ~ pop18 + totalinc + region + pop_density + docs_density + beds_density + hsgrad, data = crime_df)
summary(fit_6)
# compare nested (small vs large) models
# Ho: smaller model is defensible
anova(fit_6, fit_no_inte)
```

Reject H0, large model is superior


### no totalinc
```{r}
fit_7 = lm(crm_1000 ~ pop18 + poverty + hsgrad + beds_density + docs_density + region + pop_density, data = crime_df)
summary(fit_7)
# compare nested (small vs large) models
# Ho: smaller model is defensible
anova(fit_7, fit_no_inte)
```

Reject H0, large model is superior.

### no pop18
```{r}
fit_8 = lm(crm_1000 ~ poverty + hsgrad + beds_density + docs_density + region + pop_density + totalinc, data = crime_df)
summary(fit_7)
# compare nested (small vs large) models
# Ho: smaller model is defensible
anova(fit_8, fit_no_inte)
```

Reject H0, large model is superior.

Thus, the fit model would be: `model_selective = lm(crm_1000 ~ pop18 + totalinc + poverty + region + pop_density + docs_density + beds_density, data = crime_df)`

or `model_selective0 = lm(crm_1000 ~ pop18 + totalinc + poverty + region + pop_density + hsgrad + beds_density = crime_df)`

```{r}
mod_1 = lm(crm_1000 ~ pop18 + totalinc + poverty + region + pop_density + beds_density + docs_density, data = crime_df)
mod_2 = lm(crm_1000 ~ pop18 + totalinc + poverty + region + pop_density + beds_density, data = crime_df)


# compare nested (small vs large) models
# Ho: smaller model is defensible
anova(mod_2, mod_1)
```

Fail to reject H0, large model is not superior, we should keep the smaller one.

```{r}
mod_3 = lm(crm_1000 ~ pop18 + totalinc + poverty + region + pop_density+ beds_density + hsgrad, data = crime_df)
mod_4 = lm(crm_1000 ~ pop18 + totalinc + poverty + region + pop_density + beds_density, data = crime_df)


# compare nested (small vs large) models
# Ho: smaller model is defensible
anova(mod_4, mod_3)
```

Fail to reject H0, large model is not superior, we should keep the smaller one.

Hence, the model should be: `lm(crm_1000 ~ pop18 + totalinc + poverty + region + pop_density + beds_density, data = crime_df)`

## Potential Interaction

```{r}
coplot(crm_1000 ~ docs_density | totalinc, data = crime_df,
       number = 4, rows = 1,
       panel = panel.smooth)

coplot(crm_1000 ~ docs_density | beds_density, data = crime_df,
       number = 4, rows = 1,
       panel = panel.smooth)

coplot(crm_1000 ~ pop18 | region, data = crime_df,
       number = 4, rows = 1,
       panel = panel.smooth)

coplot(crm_1000 ~ bagrad | poverty, data = crime_df,
       number = 4, rows = 1,
       panel = panel.smooth)

coplot(crm_1000 ~ poverty | region, data = crime_df,
       number = 4, rows = 1,
       panel = panel.smooth)

coplot(crm_1000 ~ pcincome | region, data = crime_df,
       number = 4, rows = 1,
       panel = panel.smooth)
```

### model_select

Build model with interaction
```{r}
select_1 = lm(crm_1000 ~ totalinc + poverty + hsgrad + beds_density + docs_density + region + pop_density + pop18*region + poverty*region + docs_density*totalinc + docs_density*beds_density, data = crime_df)

summary(select_1)
```

Test if three interaction term is significant

```{r}
select_2 = lm(crm_1000 ~ totalinc + poverty + hsgrad + beds_density + docs_density + region + pop_density + pop18*region + poverty*region + docs_density*totalinc, data = crime_df)

anova(select_2, select_1)
```

Reject H0

```{r}
select_3 = lm(crm_1000 ~ totalinc + poverty + hsgrad + beds_density + docs_density + region + pop_density + pop18*region + poverty*region + docs_density*beds_density, data = crime_df)

anova(select_3, select_1)
```

Reject H0

```{r}
select_4 = lm(crm_1000 ~ totalinc + poverty + hsgrad + beds_density + docs_density + region + pop_density + pop18*region + docs_density*totalinc + docs_density*beds_density, data = crime_df)

anova(select_4, select_1)
```

Accept H0

```{r}
select_5 = lm(crm_1000 ~ totalinc + poverty + hsgrad + beds_density + docs_density + region + pop_density + docs_density*totalinc + docs_density*beds_density, data = crime_df)

anova(select_5, select_1)
```

Accept H0

### model_backward
```{r}
backward_1 = lm(crm_1000 ~ pop18 + hsgrad + bagrad + poverty + pcincome + 
    totalinc + region + pop_density + beds_density + pop18*region + bagrad*poverty + poverty*region + pcincome*region, data = crime_df)
summary(backward_1)
```

Test if interaction term is significant

```{r}
backward_2 = lm(crm_1000 ~ pop18 + hsgrad + bagrad + poverty + pcincome + 
    totalinc + region + pop_density + beds_density + bagrad*poverty + poverty*region + pcincome*region, data = crime_df)

anova(backward_2, backward_1)
```

Not Reject H0

```{r}
backward_3 = lm(crm_1000 ~ pop18 + hsgrad + bagrad + poverty + pcincome + 
    totalinc + region + pop_density + beds_density + poverty*region + pcincome*region, data = crime_df)

anova(backward_3, backward_1)
```

Reject H0

```{r}
backward_4 = lm(crm_1000 ~ pop18 + hsgrad + bagrad + poverty + pcincome + 
    totalinc + region + pop_density + beds_density + bagrad*poverty + pcincome*region, data = crime_df)
anova(backward_4, backward_1)
```

Reject H0

```{r}
backward_5 = lm(crm_1000 ~ pop18 + hsgrad + bagrad + poverty + pcincome + 
    totalinc + region + pop_density + beds_density + bagrad*poverty + poverty*region, data = crime_df)

anova(backward_5, backward_1)
```

Accept H0

## Candidate models

Thus the four candidate regression model should be:
```{r}
model_1 = lm(crm_1000 ~ totalinc + poverty + hsgrad + beds_density + docs_density + region + pop_density + docs_density*totalinc + docs_density*beds_density, data = crime_df)

model_2 = lm(crm_1000 ~ pop18 + hsgrad + bagrad + poverty + pcincome + 
    totalinc + region + pop_density + beds_density + bagrad*poverty + poverty*region, data = crime_df)

model_3 = lm(crm_1000 ~ pop18 + hsgrad + bagrad + poverty + pcincome + 
    totalinc + region + pop_density + beds_density + bagrad*poverty, data = crime_df)
```

## Diagnostic: Boxcox
```{r}
#model 1
bc <- boxcox(model_1, lambda = seq(-5, 5, by = 0.25))
lambda <- bc$x[which.max(bc$y)]
#choose power 1/2

crime_df$crm.bc <- crime_df$crm_1000^(1/2)
model_bc1 = lm(crm.bc ~ totalinc + poverty + hsgrad + beds_density + docs_density + region + pop_density + docs_density*totalinc + docs_density*beds_density, data = crime_df)

par(mfrow = c(1,2))
plot(model_1, which = 2)
plot(model_bc1, which = 2)

summary(model_1)
summary(model_bc1)
```

```{r}
#model 2
bc <- boxcox(model_2, lambda = seq(-5, 5, by = 0.25))
lambda <- bc$x[which.max(bc$y)]
#choose power 1/2

crime_df$crm.bc <- crime_df$crm_1000^(1/2)
model_bc2 = lm(crm.bc ~ pop18 + hsgrad + bagrad + poverty + pcincome + 
    totalinc + region + pop_density + beds_density + bagrad*poverty + poverty*region, data = crime_df)

par(mfrow = c(1,2))
plot(model_2, which = 2)
plot(model_bc2, which = 2)

summary(model_2)
summary(model_bc2)
```

```{r}
#model 3
bc <- boxcox(model_3, lambda = seq(-5, 5, by = 0.25))
lambda <- bc$x[which.max(bc$y)]
#choose power 1/2

crime_df$crm.bc <- crime_df$crm_1000^(1/2)
model_bc3 = lm(crm.bc ~ pop18 + hsgrad + bagrad + poverty + pcincome + 
    totalinc + region + pop_density + beds_density + bagrad*poverty, data = crime_df)

par(mfrow = c(1,2))
plot(model_3, which = 2)
plot(model_bc3, which = 2)

summary(model_3)
summary(model_bc3)
```