---
title: "Partial ANOVA and Interaction"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include = FALSE, message = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

set.seed(1)
```

## Data upload

```{r, message = FALSE}
crime_df = 
  read_csv("cdi.csv") %>% 
  mutate(
    region = factor(region),
    pop_density = pop / area,
    crm_1000 = 1000 * crimes / pop
  ) %>% 
  select(crm_1000, everything())
```

## Partial ANOVA with predictors

```{r}
#linear regression model with full predictors
fit_no_inte = lm(crm_1000 ~ totalinc + poverty + hsgrad + beds + docs + region + pop_density, data = crime_df)

summary(fit_no_inte)
```

Next using partial ANOVA to test whether large model is superior

```{r}
# no pop_density
fit_1 = lm(crm_1000 ~ totalinc + poverty + hsgrad + beds + docs + region, data = crime_df)
summary(fit_1)

# compare nested (small vs large) models
# Ho: smaller model is defensible
anova(fit_1, fit_no_inte)


```

Reject H0, large model is superior

```{r}
# no region
fit_no_inte = lm(crm_1000 ~ totalinc + poverty + hsgrad + beds + docs + pop_density + region , data = crime_df)
fit_2 = lm(crm_1000 ~ totalinc + poverty + hsgrad + beds + docs + pop_density, data = crime_df)
summary(fit_2)

# compare nested (small vs large) models
# Ho: smaller model is defensible
anova(fit_2, fit_no_inte)


```

Reject H0, large model is superior


```{r}
# no docs
fit_no_inte = lm(crm_1000 ~ totalinc + poverty + hsgrad + beds + region + pop_density + docs, data = crime_df)
fit_3 = lm(crm_1000 ~ totalinc + poverty + hsgrad + beds + region + pop_density, data = crime_df)
summary(fit_3)

# compare nested (small vs large) models
# Ho: smaller model is defensible
anova(fit_3, fit_no_inte)


```

Reject H0, large model is superior

```{r}
# no beds
fit_no_inte = lm(crm_1000 ~ totalinc + poverty + hsgrad + region + pop_density + docs+ beds, data = crime_df)
fit_4 = lm(crm_1000 ~ totalinc + poverty + hsgrad + region + pop_density + docs, data = crime_df)
summary(fit_4)

# compare nested (small vs large) models
# Ho: smaller model is defensible
anova(fit_4, fit_no_inte)


```

Reject H0, large model is superior

```{r}
# no hsgrad
fit_no_inte = lm(crm_1000 ~ totalinc + poverty + region + pop_density + docs+ beds + hsgrad , data = crime_df)
fit_5 = lm(crm_1000 ~ totalinc + poverty + region + pop_density + docs + beds, data = crime_df)
summary(fit_5)

# compare nested (small vs large) models
# Ho: smaller model is defensible
anova(fit_5, fit_no_inte)


```

Reject H0, large model is superior

```{r}
# no poverty
fit_no_inte = lm(crm_1000 ~ totalinc + region + pop_density + docs+ beds + hsgrad + poverty , data = crime_df)
fit_6 = lm(crm_1000 ~ totalinc + region + pop_density + docs + beds + hsgrad, data = crime_df)
summary(fit_6)

# compare nested (small vs large) models
# Ho: smaller model is defensible
anova(fit_6, fit_no_inte)


```

Reject H0, large model is superior

```{r}
# no totalinc
fit_no_inte = lm(crm_1000 ~ poverty + hsgrad + beds + docs + region + pop_density + totalinc, data = crime_df)
fit_7 = lm(crm_1000 ~ poverty + hsgrad + beds + docs + region + pop_density, data = crime_df)
summary(fit_7)

# compare nested (small vs large) models
# Ho: smaller model is defensible
anova(fit_7, fit_no_inte)


```

Fail to reject H0, large model is not superior, then we do not include total person income.


## Potential Interaction

* Number of active physicians & Number of hospital beds; 
* Number of hospital beds & Total personal income; 
* Number of active physicians & Total personal income; 

Consider the variable "total person income" to be the confounder of beds and docs variables?

Build model with interaction
```{r}
inter_initial = lm(crm_1000 ~ poverty + hsgrad + beds + docs + region + pop_density + beds * docs + docs * totalinc + beds * totalinc, data = crime_df)

summary(inter_initial)
```

### Test if three interaction term is significant

```{r}
# beds and docs
bed_doc = lm(crm_1000 ~ poverty + hsgrad + beds + docs + region + pop_density + beds * totalinc + docs * totalinc, data = crime_df)

anova(bed_doc, inter_initial)
```

Reject H0


```{r}
# beds and totalinc
bed_totalinc = lm(crm_1000 ~ poverty + hsgrad + beds + docs + region + pop_density + beds * docs + docs * totalinc, data = crime_df)

anova(bed_totalinc, inter_initial)
```

Fail to reject H0, use the simple regression model


```{r}
# docs and totalinc
doc_totalinc = lm(crm_1000 ~ poverty + hsgrad + beds + docs + region + pop_density + beds * docs + beds * totalinc, data = crime_df)

anova(doc_totalinc, inter_initial)
```

Fail to reject H0, use the simple regression model

Thus the two candidate regression model should be:
```{r}
doc_totalinc = lm(crm_1000 ~ poverty + hsgrad + beds + docs + region + pop_density + beds * docs + beds * totalinc, data = crime_df)
broom::tidy(doc_totalinc)

```


```{r}
bed_totalinc = lm(crm_1000 ~ poverty + hsgrad + beds + docs + region + pop_density + beds * docs + docs * totalinc, data = crime_df)
broom::tidy(bed_totalinc)
```





